---
title: "post_treatment_veg"
format: html
editor: visual
---

## reading in packages & data

```{r}
#| message: false
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)
library(car)
library(effectsize)
library(rstatix)
library(multcomp)

veg_data <- read_xlsx(path = here("data", 
                        "post_treatment_veg_all.xlsx"),
                         sheet = "CCBER_VegMonitoring_Quadrat_4")
```

## veg data cleaning
```{r}
veg_data_clean <- veg_data %>% 
  clean_names() %>% 
# select(global_id, monitoring_date, unlisted_lagoon_transect_name_number, transect_distance_of_quadrat, count_of_native_species, sum_of_native_cover, sum_of_non_native_cover, count_of_non_native_species, sum_of_unknown_cover, sum_of_thatch, bare_ground, sum_of_all_cover) %>% 
  rename(treatment = unlisted_lagoon_transect_name_number,
         native_count = count_of_native_species,
         native_cover = sum_of_native_cover,
         nonnative_count = count_of_non_native_species,
         nonnative_cover = sum_of_non_native_cover,
         thatch_cover = sum_of_thatch) %>% 
  mutate(total_cover = sum_of_all_cover - sum_of_unknown_cover) %>% 
  mutate(weeding = case_when(
             treatment == "Thatch-None" ~ "Thatch Removal",
             treatment == "Thatch-Low" ~ "Thatch Removal",
             treatment == "Thatch-High" ~ "Thatch Removal",
             treatment == "Control-None" ~ "Control",
             treatment == "Control-Low" ~ "Control",
             treatment == "Control-High" ~ "Control",
             treatment == "Burn-None" ~ "Burn",
             treatment == "Burn-Low" ~ "Burn",
             treatment == "Burn-High" ~ "Burn",
             treatment == "Clethodim-None" ~ "Herbicide",
             treatment == "Clethodim-Low" ~ "Herbicide",
             treatment == "Clethodim-High" ~ "Herbicide"),
         seeding = case_when(
             treatment == "Thatch-None" ~ "None",
             treatment == "Thatch-Low" ~ "Low",
             treatment == "Thatch-High" ~ "High",
             treatment == "Control-None" ~ "None",
             treatment == "Control-Low" ~ "Low",
             treatment == "Control-High" ~ "High",
             treatment == "Burn-None" ~ "None",
             treatment == "Burn-Low" ~ "Low",
             treatment == "Burn-High" ~ "High",
             treatment == "Clethodim-None" ~ "None",
             treatment == "Clethodim-Low" ~ "Low",
             treatment == "Clethodim-High" ~ "High")) %>% 
  mutate(weeding= fct_relevel(weeding, "Control", "Thatch Removal", "Herbicide", "Burn"),
         seeding = fct_relevel(seeding, "None", "Low", "High"))
```

## cover/density weeding plots
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = native_cover, color = weeding) +
  geom_boxplot() +
 geom_point(position=position_jitterdodge(
   jitter.width = 0,
   jitter.height = 0,
    seed = NULL),
            alpha = 0.3) +
  theme_minimal() +
  ylim(0,120) +
  scale_color_manual(values = c("#38A800", "#FF327C", "#00A9E6", "#F77600")) +
  labs(x = "Treatment",
       y = "Cover (%)",
       color = "Treatment")
       #title = "Non-Native Density by Weeding Treatment") 
```

## cover/density seeding or seeding/weeding plots
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = native_count, color = seeding) +
  geom_boxplot() +
 geom_point(position=position_jitterdodge(
   jitter.width = 0,
   jitter.height = 0,
    seed = NULL),
            alpha = 0.3) +
  theme_minimal() +
  scale_color_manual(values = c("#FFC567", "#67C0AC", "#B660CD")) +
  labs(x = "Weeding Treatment",
       y = "Count",
       color = "Seeding Rate",
       title = "Native Density by Weeding and Seeding Treatments")
```

# using * for interaction, but do we care? should i do + instead?

# stats for native cover - OKAY???
```{r}
## normal dist
hist(veg_data_clean $ native_cover)

nativecover_aov <- aov(sqrt(native_cover) ~ weeding * seeding, data = veg_data_clean)

summary(nativecover_aov)

plot(nativecover_aov)

cooksnativeaov <- cooks.distance(nativecover_aov)

plot(cooksnativeaov)

# this is fairly normal, now testing for equal variances

leveneTest(sqrt(native_cover) ~ weeding * seeding, veg_data_clean)

# significant, variances are not equal, p = 0.02

oneway.test(native_cover ~  weeding * seeding,
  data = veg_data_clean,
  var.equal = FALSE) # assuming unequal variances

## trying to do a posthoc for welch?????
games_howell_test(data = veg_data_clean, 
                  formula = native_cover ~  weeding * seeding, 
                  conf.level = 0.95, 
                  detailed = FALSE) 
# that does not work ^

## if i proceed with anova anyway -->

TukeyHSD(nativecover_aov)

## effect size
effectsize::eta_squared(nativecover_aov)

```

# anova stats for nonnative cover - OKAY???
```{r}
## normal dist
hist(veg_data_clean $ nonnative_cover)

nonnativecover_aov <- aov(sqrt(nonnative_cover) ~ weeding * seeding, data = veg_data_clean)

plot(nonnativecover_aov)

summary(nonnativecover_aov)

cooksnonnativeaov <- cooks.distance(nonnativecover_aov)

plot(cooksnonnativeaov)

## seems normal-ish??? worried about scale-location plot, but qq and constant leverage seem okay

leveneTest(sqrt(nonnative_cover) ~ weeding * seeding, veg_data_clean)

# significant, variances are not equal, p = 0.0009

## effect size
effectsize::eta_squared(nonnativecover_aov)

## post-hoc Tukey
TukeyHSD(nonnativecover_aov)

### IS THIS WRONG TO DO WITH WEEDING AND SEEDING VARIABLES? THERE ARE SO MANY OUTPUTS? DO I JUST FIND EACH SIGNIFICANT ONE MANUALLY?

```

# kruskal wallis for nonative cover
```{r}
kruskal.test(nonnative_cover ~ seeding, # formula
             data = veg_data_clean) # data

# seeding - 

#can I do kruskal wallis with two variables?

dunn_test(nonnative_cover ~ seeding, # formula
          data = veg_data_clean) # data

kruskal_effsize(nonnative_cover ~ weeding, # formula
                data = veg_data_clean) # data

# herbicide and burn have significantly different nonnative covers compared to control and thatch-removal, but not from each other, large effect size of 0.74. seeding has no significant effect on nn cover 

# We found a difference in nonnative cover between weeding treatments (insert KW info here). There was a large effect of weeding on nonnative cover (n^2 = 0.74), with burn and herbicide reducing nonnative cover more than control and thatch-removal (insert Dunn’s post-hoc info here).
```

# stats for native density - GOOD
```{r}
hist(veg_data_clean $ native_count)

nativedensity_aov <- aov(native_count ~ weeding * seeding, data = veg_data_clean)

summary(nativedensity_aov)

plot(nativedensity_aov)

# looks normal! scale-location is  a bit wonky but mostly okay

leveneTest(native_count ~ weeding * seeding, veg_data_clean)

# Variances are equal!

# I can use my ANOVA

## Tukey-HSD
TukeyHSD(nativedensity_aov)

# ^ does this makes sense to do when it gives back EACH combination, or is there a way to group to compare just weeding and just seeding to each other. I ended up looking at the ANOVA p stats for weeding and seeding sep?

## effect size
effectsize::eta_squared(nativedensity_aov)


# We found a significant, large effect (η2 = 0.73) of weeding treatments (F(84, 3) = 77.10, p < 0.001, ⍺ = 0.05) on native density 

# There was a significant, small effect (η2 = 0.15) of seeding treatments (F(84, 2) = 7.13, p = 0.001, , ⍺ = 0.05) on native density.

# The controlled burn (M = appears to have the greatest effect in increasing native density
# No seeding appears to have the least effect in increasing native density. There does not appear to be a significant difference between high and low seeding rate treatments, indicating that any amount of native seed applied to the landscape is important in increasing native density.
```

# stats for nonnative density - GOOD
```{r}
hist(veg_data_clean $ nonnative_count)

nonnativedensity_aov <- aov(nonnative_count ~ weeding * seeding, data = veg_data_clean)

summary(nonnativedensity_aov)

plot(nonnativedensity_aov)

# SOOOO NORMAL

leveneTest(native_count ~ weeding * seeding, veg_data_clean)

# Variances are equal!

# I can use my ANOVA!

## Tukey-HSD
TukeyHSD(nonnativedensity_aov)

## effect size
effectsize::eta_squared(nonnativedensity_aov)

```






## native cover and weeding statistical tests
```{r}

## normal distribution
ggplot(veg_data_clean) +
  aes(x = native_cover) +
  geom_histogram(bins = 9) +
  facet_wrap(~weeding,
             scales = "free")

## qqplot
ggplot(data = veg_data_clean,
       aes(sample = native_cover)) +
  geom_qq_line() +
  geom_qq() +
  facet_wrap(~weeding,
             scales = "free")

## shapiro-wilks
burn_native_sw <- veg_data_clean %>% 
  filter(weeding == "Burn") %>% 
  pull(native_cover)

shapiro.test(burn_native_sw)

## equal variances
veg_data_clean %>% 
  group_by(weeding) %>% 
  reframe (var = var(native_cover))

leveneTest(native_cover ~ weeding,
           data = veg_data_clean)

## ANOVA
native_weeding_anova <- aov(native_cover ~ weeding * seeding,
                            data = veg_data_clean)

summary(native_weeding_anova)

## Tukey-HSD
TukeyHSD(native_weeding_anova)

## effect size
effectsize::eta_squared(native_weeding_anova)





# We found a large effect (η2 = 0.63) of weeding treatments on native cover (one-way ANOVA, F(3, 92) = 52.99, p < 0.001, ⍺ = 0.05). On average, native cover in the burn plot was higher than cover in the herbicide (p < 0.001), thatch removal (p < 0.001), and control (p< 0.001) plots.
```

```{r}
hist(veg_data_clean $ native_cover) ##kruskal-wallis & dunns test instead of anova for nonnative

report(native_weeding_anova)
```


## nonnative cover and weeding plot
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = nonnative_cover, color = weeding) +
  geom_boxplot() +
  geom_jitter(width = 0,
              height = 0,
              alpha = 0.3) +
  theme_minimal() +
  scale_color_manual(values = c("#38A800", "#FF327C", "#00A9E6", "#F77600")) +
  labs(x = "Weeding",
       y = "Non-native Cover",
       color = "Weeding")
```

## nonnative cover and weeding statistical tests
```{r}
## normal distribution
ggplot(veg_data_clean) +
  aes(x = nonnative_cover) +
  geom_histogram(bins = 9) +
  facet_wrap(~weeding,
             scales = "free")

## qqplot
ggplot(data = veg_data_clean,
       aes(sample = nonnative_cover)) +
  geom_qq_line() +
  geom_qq() +
  facet_wrap(~weeding,
             scales = "free")

## shapiro-wilks
burn_nonnative_sw <- veg_data_clean %>% 
  filter(weeding == "Burn") %>% 
  pull(nonnative_cover)

shapiro.test(burn_nonnative_sw)

## equal variances
veg_data_clean %>% 
  group_by(weeding) %>% 
  reframe (var = var(nonnative_cover))

leveneTest(nonnative_cover ~ weeding,
           data = veg_data_clean)

## ANOVA
nonnative_weeding_anova <- aov(nonnative_cover ~ weeding * seeding,
                            data = veg_data_clean)

summary(nonnative_weeding_anova)

## Tukey-HSD
TukeyHSD(nonnative_weeding_anova)

## effect size
effectsize::eta_squared(nonnative_weeding_anova)

# We found a large effect (η2 = 0.47) of weeding treatments on native cover (one-way ANOVA, F(3, 92) = 27.36, p < 0.001, ⍺ = 0.05). On average, non-native cover in the burn and herbicide plots was lower than cover in thatch removal (herbicide: p < 0.001, burn: p < 0.001) and control (herbicide: p <0.001, burn: p < 0.001) plots.
```



## native density and weeding plot
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = native_count, color = weeding) +
  geom_boxplot() +
  geom_jitter(width = 0,
              height = 0,
              alpha = 0.3) +
  theme_minimal() +
  scale_color_manual(values = c("#38A800", "#FF327C", "#00A9E6", "#F77600")) +
  labs(x = "Weeding",
       y = "Native Density",
       color = "Weeding")
```

## native density and weeding statistical tests
```{r}
## normal distribution
ggplot(veg_data_clean) +
  aes(x = native_count) +
  geom_histogram(bins = 9) +
  facet_wrap(~weeding,
             scales = "free")

## qqplot
ggplot(data = veg_data_clean,
       aes(sample = native_count)) +
  geom_qq_line() +
  geom_qq() +
  facet_wrap(~weeding,
             scales = "free")

## shapiro-wilks
burn_native_density_sw <- veg_data_clean %>% 
  filter(weeding == "Burn") %>% 
  pull(native_count)

shapiro.test(burn_native_density_sw)

## equal variances
veg_data_clean %>% 
  group_by(weeding) %>% 
  reframe (var = var(native_count))

leveneTest(native_count ~ weeding,
           data = veg_data_clean)

## ANOVA
native_weeding_density_anova <- aov(native_count ~ weeding,
                            data = veg_data_clean)

summary(native_weeding_density_anova)

## Tukey-HSD
TukeyHSD(native_weeding_density_anova)

## effect size
effectsize::eta_squared(native_weeding_density_anova)

# We found a large effect (η2 = 0.63) of weeding treatments on native cover (one-way ANOVA, F(3, 92) = 52.99, p < 0.001, ⍺ = 0.05). On average, native cover in the burn plot was higher than cover in the herbicide (p < 0.001), thatch removal (p < 0.001), and control (p< 0.001) plots.
```


## nonnative density and weeding plot
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = nonnative_count, color = weeding) +
  geom_boxplot() +
  geom_jitter(width = 0,
              height = 0,
              alpha = 0.3) +
  theme_minimal() +
  scale_color_manual(values = c("#38A800", "#FF327C", "#00A9E6", "#F77600")) +
  labs(x = "Weeding",
       y = "Non-native Density",
       color = "Weeding")
```

## nonnative density and weeding statistical tests
```{r}
## normal distribution
ggplot(veg_data_clean) +
  aes(x = nonnative_count) +
  geom_histogram(bins = 9) +
  facet_wrap(~weeding,
             scales = "free")

## qqplot
ggplot(data = veg_data_clean,
       aes(sample = nonnative_count)) +
  geom_qq_line() +
  geom_qq() +
  facet_wrap(~weeding,
             scales = "free")

## shapiro-wilks
burn_nonnative_density_sw <- veg_data_clean %>% 
  filter(weeding == "Burn") %>% 
  pull(nonnative_count)

shapiro.test(burn_nonnative_density_sw)

## equal variances
veg_data_clean %>% 
  group_by(weeding) %>% 
  reframe (var = var(nonnative_count))

leveneTest(nonnative_count ~ weeding,
           data = veg_data_clean)

## ANOVA
nonnative_weeding_density_anova <- aov(nonnative_count ~ weeding,
                            data = veg_data_clean)

summary(nonnative_weeding_density_anova)

## Tukey-HSD
TukeyHSD(nonnative_weeding_density_anova)

## effect size
effectsize::eta_squared(nonnative_weeding_density_anova)

# We found a large effect (η2 = 0.47) of weeding treatments on native cover (one-way ANOVA, F(3, 92) = 27.36, p < 0.001, ⍺ = 0.05). On average, non-native cover in the burn and herbicide plots was lower than cover in thatch removal (herbicide: p < 0.001, burn: p < 0.001) and control (herbicide: p <0.001, burn: p < 0.001) plots.
```




## thatch cover & natives
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = thatch_cover, color = weeding) +
 geom_boxplot() +
 geom_point(position=position_jitterdodge(
   jitter.width = 0,
   jitter.height = 0,
    seed = NULL),
            alpha = 0.3) +
  theme_minimal() +
  ylim(0,100) +
  scale_color_manual(values = c("#38A800", "#FF327C", "#00A9E6", "#F77600")) +
  labs(x = "Treatment",
       y = "Cover (%)",
       color = "Treatment")

ggplot(veg_data_clean) +
  aes(x = thatch_cover, y = nonnative_cover, color = weeding) +
  geom_point()

## don't include burn in thatch - nonnativecover

## bring in pre-treatment data to put side by side --> look at thatch cover change i.e. how effective 
## bring in fire temp for burn cover within blocks
```


## boxplots & outliers
```{r}
boxplot(veg_data_clean$native_cover ~ veg_data_clean$weeding)
```
## anova
```{r}
res_aov <- aov(native_cover ~ weeding,
              data = veg_data_clean)

hist(res_aov$residuals)


```

