---
title: "post_treatment_veg"
format: html
editor: visual
---

## reading in packages & data

```{r}
#| message: false
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)
library(car)
library(effectsize)
library(rstatix)

veg_data <- read_xlsx(path = here("data", 
                        "post_treatment_veg_all.xlsx"),
                         sheet = "CCBER_VegMonitoring_Quadrat_4")

burn_temps <- read_xlsx(path = here("data", 
                        "Burn Temps.xlsx"))
```

## veg data cleaning
```{r}
veg_data_clean <- veg_data %>% 
  clean_names() %>% 
  select(global_id, monitoring_date, unlisted_lagoon_transect_name_number, transect_distance_of_quadrat, count_of_native_species, sum_of_native_cover, sum_of_non_native_cover, count_of_non_native_species, sum_of_unknown_cover, sum_of_thatch, bare_ground, sum_of_all_cover) %>% 
  rename(treatment = unlisted_lagoon_transect_name_number,
         native_count = count_of_native_species,
         native_cover = sum_of_native_cover,
         nonnative_count = count_of_non_native_species,
         nonnative_cover = sum_of_non_native_cover,
         thatch_cover = sum_of_thatch) %>% 
  mutate(total_cover = sum_of_all_cover - sum_of_unknown_cover) %>% 
  mutate(weeding = case_when(
             treatment == "Thatch-None" ~ "Thatch Removal",
             treatment == "Thatch-Low" ~ "Thatch Removal",
             treatment == "Thatch-High" ~ "Thatch Removal",
             treatment == "Control-None" ~ "Control",
             treatment == "Control-Low" ~ "Control",
             treatment == "Control-High" ~ "Control",
             treatment == "Burn-None" ~ "Burn",
             treatment == "Burn-Low" ~ "Burn",
             treatment == "Burn-High" ~ "Burn",
             treatment == "Clethodim-None" ~ "Herbicide",
             treatment == "Clethodim-Low" ~ "Herbicide",
             treatment == "Clethodim-High" ~ "Herbicide"),
         seeding = case_when(
             treatment == "Thatch-None" ~ "None",
             treatment == "Thatch-Low" ~ "Low",
             treatment == "Thatch-High" ~ "High",
             treatment == "Control-None" ~ "None",
             treatment == "Control-Low" ~ "Low",
             treatment == "Control-High" ~ "High",
             treatment == "Burn-None" ~ "None",
             treatment == "Burn-Low" ~ "Low",
             treatment == "Burn-High" ~ "High",
             treatment == "Clethodim-None" ~ "None",
             treatment == "Clethodim-Low" ~ "Low",
             treatment == "Clethodim-High" ~ "High")) %>% 
  mutate(weeding= fct_relevel(weeding, "Control", "Thatch Removal", "Herbicide", "Burn"))
```

## burn dc
```{r}
# cleaning up data
burn_temps_clean <- burn_temps %>% clean_names() %>%
  rename(distance = distance_from_edge_m)

burn_temps_avg <- burn_temps_clean %>%
  group_by(distance, edge, depth) %>% 
  reframe(mean = mean(max_temp)) %>% 
  mutate(depth = fct_relevel(depth, "surface", "sub-surface"))
```


## native cover and weeding plot
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = native_cover, color = weeding) +
  geom_boxplot() +
  geom_jitter(width = 0,
              height = 0,
              alpha = 0.3) +
  theme_minimal() +
  scale_color_manual(values = c("#38A800", "#FF327C", "#00A9E6", "#F77600")) +
  labs(x = "Weeding",
       y = "Native Cover",
       color = "Weeding")
```

## native cover and weeding statistical tests
```{r}
## normal distribution
ggplot(veg_data_clean) +
  aes(x = native_cover) +
  geom_histogram(bins = 9) +
  facet_wrap(~weeding,
             scales = "free")

## qqplot
ggplot(data = veg_data_clean,
       aes(sample = native_cover)) +
  geom_qq_line() +
  geom_qq() +
  facet_wrap(~weeding,
             scales = "free")

## shapiro-wilks
burn_native_sw <- veg_data_clean %>% 
  filter(weeding == "Burn") %>% 
  pull(native_cover)

shapiro.test(burn_native_sw)

## equal variances
veg_data_clean %>% 
  group_by(weeding) %>% 
  reframe (var = var(native_cover))

leveneTest(native_cover ~ weeding,
           data = veg_data_clean)

## ANOVA
native_weeding_anova <- aov(native_cover ~ weeding,
                            data = veg_data_clean)

summary(native_weeding_anova)

## Tukey-HSD
TukeyHSD(native_weeding_anova)

## effect size
effectsize::eta_squared(native_weeding_anova)

# We found a large effect (η2 = 0.63) of weeding treatments on native cover (one-way ANOVA, F(3, 92) = 52.99, p < 0.001, ⍺ = 0.05). On average, native cover in the burn plot was higher than cover in the herbicide (p < 0.001), thatch removal (p < 0.001), and control (p< 0.001) plots.
```


## nonnative cover and weeding plot
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = nonnative_cover, color = weeding) +
  geom_boxplot() +
  geom_jitter(width = 0,
              height = 0,
              alpha = 0.3) +
  theme_minimal() +
  scale_color_manual(values = c("#38A800", "#FF327C", "#00A9E6", "#F77600")) +
  labs(x = "Weeding",
       y = "Non-native Cover",
       color = "Weeding")
```

## nonnative cover and weeding statistical tests
```{r}
## normal distribution
ggplot(veg_data_clean) +
  aes(x = nonnative_cover) +
  geom_histogram(bins = 9) +
  facet_wrap(~weeding,
             scales = "free")

## qqplot
ggplot(data = veg_data_clean,
       aes(sample = nonnative_cover)) +
  geom_qq_line() +
  geom_qq() +
  facet_wrap(~weeding,
             scales = "free")

## shapiro-wilks
burn_nonnative_sw <- veg_data_clean %>% 
  filter(weeding == "Burn") %>% 
  pull(nonnative_cover)

shapiro.test(burn_nonnative_sw)

## equal variances
veg_data_clean %>% 
  group_by(weeding) %>% 
  reframe (var = var(nonnative_cover))

leveneTest(nonnative_cover ~ weeding,
           data = veg_data_clean)

## ANOVA
nonnative_weeding_anova <- aov(nonnative_cover ~ weeding,
                            data = veg_data_clean)

summary(nonnative_weeding_anova)

## Tukey-HSD
TukeyHSD(nonnative_weeding_anova)

## effect size
effectsize::eta_squared(nonnative_weeding_anova)

# We found a large effect (η2 = 0.47) of weeding treatments on native cover (one-way ANOVA, F(3, 92) = 27.36, p < 0.001, ⍺ = 0.05). On average, non-native cover in the burn and herbicide plots was lower than cover in thatch removal (herbicide: p < 0.001, burn: p < 0.001) and control (herbicide: p <0.001, burn: p < 0.001) plots.
```



## native density and weeding plot
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = native_count, color = weeding) +
  geom_boxplot() +
  geom_jitter(width = 0,
              height = 0,
              alpha = 0.3) +
  theme_minimal() +
  scale_color_manual(values = c("#38A800", "#FF327C", "#00A9E6", "#F77600")) +
  labs(x = "Weeding",
       y = "Native Density",
       color = "Weeding")
```

## native density and weeding statistical tests
```{r}
## normal distribution
ggplot(veg_data_clean) +
  aes(x = native_count) +
  geom_histogram(bins = 9) +
  facet_wrap(~weeding,
             scales = "free")

## qqplot
ggplot(data = veg_data_clean,
       aes(sample = native_count)) +
  geom_qq_line() +
  geom_qq() +
  facet_wrap(~weeding,
             scales = "free")

## shapiro-wilks
burn_native_density_sw <- veg_data_clean %>% 
  filter(weeding == "Burn") %>% 
  pull(native_count)

shapiro.test(burn_native_density_sw)

## equal variances
veg_data_clean %>% 
  group_by(weeding) %>% 
  reframe (var = var(native_count))

leveneTest(native_count ~ weeding,
           data = veg_data_clean)

## ANOVA
native_weeding_density_anova <- aov(native_count ~ weeding,
                            data = veg_data_clean)

summary(native_weeding_density_anova)

## Tukey-HSD
TukeyHSD(native_weeding_density_anova)

## effect size
effectsize::eta_squared(native_weeding_density_anova)

# We found a large effect (η2 = 0.63) of weeding treatments on native cover (one-way ANOVA, F(3, 92) = 52.99, p < 0.001, ⍺ = 0.05). On average, native cover in the burn plot was higher than cover in the herbicide (p < 0.001), thatch removal (p < 0.001), and control (p< 0.001) plots.
```


## nonnative density and weeding plot
```{r}
ggplot(veg_data_clean) +
  aes(x = weeding, y = nonnative_count, color = weeding) +
  geom_boxplot() +
  geom_jitter(width = 0,
              height = 0,
              alpha = 0.3) +
  theme_minimal() +
  scale_color_manual(values = c("#38A800", "#FF327C", "#00A9E6", "#F77600")) +
  labs(x = "Weeding",
       y = "Non-native Density",
       color = "Weeding")
```

## nonnative density and weeding statistical tests
```{r}
## normal distribution
ggplot(veg_data_clean) +
  aes(x = nonnative_count) +
  geom_histogram(bins = 9) +
  facet_wrap(~weeding,
             scales = "free")

## qqplot
ggplot(data = veg_data_clean,
       aes(sample = nonnative_count)) +
  geom_qq_line() +
  geom_qq() +
  facet_wrap(~weeding,
             scales = "free")

## shapiro-wilks
burn_nonnative_density_sw <- veg_data_clean %>% 
  filter(weeding == "Burn") %>% 
  pull(nonnative_count)

shapiro.test(burn_nonnative_density_sw)

## equal variances
veg_data_clean %>% 
  group_by(weeding) %>% 
  reframe (var = var(nonnative_count))

leveneTest(nonnative_count ~ weeding,
           data = veg_data_clean)

## ANOVA
nonnative_weeding_density_anova <- aov(nonnative_count ~ weeding,
                            data = veg_data_clean)

summary(nonnative_weeding_density_anova)

## Tukey-HSD
TukeyHSD(nonnative_weeding_density_anova)

## effect size
effectsize::eta_squared(nonnative_weeding_density_anova)

# We found a large effect (η2 = 0.47) of weeding treatments on native cover (one-way ANOVA, F(3, 92) = 27.36, p < 0.001, ⍺ = 0.05). On average, non-native cover in the burn and herbicide plots was lower than cover in thatch removal (herbicide: p < 0.001, burn: p < 0.001) and control (herbicide: p <0.001, burn: p < 0.001) plots.
```




## thatch cover & natives
```{r}
ggplot(veg_data_clean) +
  aes(x = thatch_cover, y = nonnative_cover, color = weeding) +
  geom_point() 

## don't include burn in thatch - nonnativecover

## bring in pre-treatment data to put side by side --> look at thatch cover change i.e. how effective 
## bring in fire temp for burn cover within blocks
```


## boxplots & outliers
```{r}
boxplot(veg_data_clean$native_cover ~ veg_data_clean$weeding)
```
## anova
```{r}
res_aov <- aov(native_cover ~ weeding,
              data = veg_data_clean)

hist(res_aov$residuals)


```

